---
title: "Proyecto 1. Modelización Espacio Temporal"
author: "Josseling Quiroz & Carlos Zuñiga"
date: "05/16/2022"
output: html_document
---
# Universidad de Costa Rica

## GF-0622 MODELIZACIÓN ESPACIO-TEMPORAL
#### Profesores: Dr. Christian Birkel / Dr.Adolfo Quesada Román


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r results='hide', warning = FALSE, message=FALSE}
library(FME)         # package used for calibration and uncertainty estimations
library(tidyverse)   # package used for data manipulation
library(gridExtra)   # package used for graphics with ggplot2
library(corrplot)
library(gplots)
library(RColorBrewer)
library(factoextra)
library(ggplot2)
library(ggpubr)
library(ggsci)
library(scales)
library (lubridate)
library(patchwork)
library (nsga2R)
library (mco)
```

```{r results='hide', warning = FALSE, message=FALSE}
# read data
library(readr)
inp <- read_csv("inp.csv", col_types = cols(`P (mm)...2` = col_number(), 
    `PET (mm)...3` = col_number(), `SM (mm)...4` = col_number(), 
    `ETR (mm)...5` = col_number(), `Qs (mm)...6` = col_number(), 
    `P (mm)...7` = col_number(), `PET (mm)...8` = col_number(), 
    `SM (mm)...9` = col_number(), `ETR (mm)...10` = col_number(), 
    `Qs (mm)...11` = col_number(), `P (mm)...12` = col_number(), 
    `PET (mm)...13` = col_number(), `SM (mm)...14` = col_number(), 
    `ETR (mm)...15` = col_number(), `Qs (mm)...16` = col_number(), 
    `P (mm)...17` = col_number(), `PET (mm)...18` = col_number(), 
    `SM (mm)...19` = col_number(), `ETR (mm)...20` = col_number(), 
    `Qs (mm)...21` = col_number(), `P (mm)...22` = col_number(), 
    `PET (mm)...23` = col_number(), `SM (mm)...24` = col_number(), 
    `ETR (mm)...25` = col_number(), `Qs (mm)...26` = col_number(), 
    MEI = col_number()))
View(inp)

# read input into dataframe
names(inp) <- c("Date","P1","PET1","Q1","ETA1","SM1","P2","PET2","Q2","ETA2","SM2","P3","PET3","Q3","ETA3","SM3","P4","PET4","Q4","ETA4","SM4","P5","PET5","Q5","ETA5","SM5","MEI")			            # define column names for direct use in model code
attach(inp,warn.conflicts = F)	
head(inp)
dim(inp) 

```
```{r results='hide', warning = FALSE, message=FALSE}

sumstat <- summary(inp)            # summary statistics
write.csv(sumstat, file="sumstat.csv")  # write output file to disc

# P-Q boxplots for comparison
b0 <- ggboxplot(P1,color="black", xlab="P1", ylab="P1 (mm/month)", ylim=c(0,1000)) 
b00 <- ggboxplot(Q1,color="blue", xlab="Q1", ylab="Q1 (mm/month)", ylim=c(0,1000)) 
b1 <- ggboxplot(P2,color="black", xlab="P2", ylab="P2 (mm/month)", ylim=c(0,1000)) 
b11 <- ggboxplot(Q2,color="blue", xlab="Q2", ylab="Q2 (mm/month)", ylim=c(0,1000)) 

plot <- ggarrange(b0, b00,
                  b1, b11,
                  ncol = 2, nrow = 2,  align = "hv", 
                  widths=c(10,10,10), heights=c(5,5,5))

ggexport(plotlist = plot, filename = "Ts_boxplot_P1.png", res = 300)

plot
```

### Visualizacion en grafico de barras de los percentiles de todas las 5 cuencas
```{r results='hide', warning = FALSE, message=FALSE}
percentiles5 <- read_csv("percentiles5.csv")

ggplot(percentiles5, aes(x=reorder(Percentiles, -`P(mm)rioesquinas`), y=`P(mm)rioesquinas`)) +
  ggtitle("Precipitacion Rio Esquinas") +
  xlab("Percentiles") + 
  ylab("Precipitacion") +  
  geom_col() +
  geom_text(aes(label=`P(mm)rioesquinas`), vjust=-0.3, size=3.5)

ggplot(percentiles5, aes(x=reorder(Percentiles, -`P(mm)riozota`), y=`P(mm)riozota`)) +
  ggtitle("Precipitacion Rio Zota") +
  xlab("Percentiles") + 
  ylab("Precipitacion") +  
  geom_col() +
  geom_text(aes(label=`P(mm)rioesquinas`), vjust=-0.3, size=3.5)

ggplot(percentiles5, aes(x=reorder(Percentiles, -`P(mm)riotempisque`), y=`P(mm)riotempisque`)) +
  ggtitle("Precipitacion Rio Tempisque") +
  xlab("Percentiles") + 
  ylab("Precipitacion") +  
  geom_col() +
  geom_text(aes(label=`P(mm)rioesquinas`), vjust=-0.3, size=3.5)

ggplot(percentiles5, aes(x=reorder(Percentiles, -`P(mm)rionaranjo`), y=`P(mm)rionaranjo`)) +
  ggtitle("Precipitacion Rio Naranjo") +
  xlab("Percentiles") + 
  ylab("Precipitacion") +  
  geom_col() +
  geom_text(aes(label=`P(mm)rioesquinas`), vjust=-0.3, size=3.5)

ggplot(percentiles5, aes(x=reorder(Percentiles, -`P(mm)riopiedras`), y=`P(mm)riopiedras`)) +
  ggtitle("Precipitacion Rio Piedras") +
  xlab("Percentiles") + 
  ylab("Precipitacion") +  
  geom_col() +
  geom_text(aes(label=`P(mm)rioesquinas`), vjust=-0.3, size=3.5)

```

### correlation analysis on full matrix
```{r results='hide', warning = FALSE, message=FALSE}
corinp <- cor (inp[,2:27], use = "pairwise.complete.obs", method="spearman")      # calculates correlation coefficients
write.csv (corinp, file="cor5cuencas.csv")    # output matrix

corrplot(corinp, method="pie")                # nice pie plot

```

### bivariate plots with histograms and coefficients
## matriz 1 sub cuencas Esquinas-Zota-Piedras
## matriz 2 sub cuencas tempisque y naranjo
```{r results='hide', warning = FALSE, message=FALSE}
panel.cor <- function(x, y, method="spearman", digits=3, cex.cor=1.2)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y, use = "pairwise.complete.obs", method=method)
  ra <- cor.test(x, y, method=method)$p.value
  txt <- round(r, digits)
  sig <- 1
  prefix <- ""
  if(ra <= 0.1) prefix <- "."
  if(ra <= 0.05) prefix <- "*"
  if(ra <= 0.01) prefix <- "**"
  if(ra <= 0.001) prefix <- "***"
  if(ra <= 0.001) sig <- 2
  color <- 2
  if(r > 0) color <- "black"
  #         color <- "gray10"
  if(r < 0) color <- "gray50"
  txt <- paste(txt, prefix, sep="\n")
  text(0.5, 0.5, txt, cex = cex.cor, font=sig, col=color)
}


```

## Put histograms on the diagonal
```{r results='hide', warning = FALSE, message=FALSE}
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 2) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
  #         rect(breaks[-nB], 0, breaks[-1], y, col="gray", ...)
}

op <- par(mfrow=c(1,1), pty="s")      ##### create correlation matrix with histograms and coefficients
pairs(inp[,2:16],
      lower.panel = panel.cor, upper.panel = panel.smooth,
      diag.panel = panel.hist)
par(op)  


panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 2) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col="cyan", ...)
  #         rect(breaks[-nB], 0, breaks[-1], y, col="gray", ...)
}

op <- par(mfrow=c(1,1), pty="s")        ##### create correlation matrix with histograms and coefficients
pairs(inp[,16:27],
      lower.panel = panel.cor, upper.panel = panel.smooth,
      diag.panel = panel.hist)

```

### Regression analysis
```{r results='hide', warning = FALSE, message=FALSE}

PQ1 <- ggscatter(inp, x = "P1", y = "Q1",xlim=c(0,1000), ylim=c(0,1000),
                 size = 4, add = "reg.line")+
  stat_cor(label.x=200, label.y = -20) +
  stat_regline_equation(label.x=200, label.y = 0)
p1 <- PQ1 + labs(x = "Rainfall P (mm/month)", y = " Streamflow (mm/month)")

PQ2 <- ggscatter(inp, x = "P2", y = "Q2",xlim=c(0,1000), ylim=c(0,1000),
                 size = 4, add = "reg.line")+
  stat_cor(label.x=200, label.y = -20) +
  stat_regline_equation(label.x=200, label.y = 0)
p2 <- PQ2 + labs(x = "Rainfall P (mm/month)", y = "Streamflow (mm/month)")

grid.arrange(p1, p2, ncol = 2, nrow=1)


```


```{r results='hide', warning = FALSE, message=FALSE}

```


```{r results='hide', warning = FALSE, message=FALSE}

```


